---
title: "Assgn4"
author: "Vaishnavi"
date: "2023-11-19"
output: html_document
---
I will use the k-implies grouping approach to conduct a non-various tiered bunch examination for this project. The objective is to divide the data into consistent groups in order to extract relevant information. Initially, we need to stack the first dataset together with the necessary bundles. It contains data from roughly twenty-one pharmaceutical companies.

The rationale behind the choices of market capitalization, leverage, rev growth, net profit margin, beta, PE ratio, ROE, and ROA

The variables that were selected are common financial measures that are employed in business performance evaluation and comparison. Market Cap, Beta, PE Ratio, ROE, ROA, Turnover of Assets, Leverage, Rev Growth, and Net Profit Margin are a few of them. When taken as a whole, these variables provide a complete view of a company's productivity, profitability, and stability.


```{r}
library(tidyverse)  # data manipulation
library(cluster)    # clustering algorithms
library(factoextra) # clustering algorithms & visualization
library(ggplot2)
library(ISLR)
library(gridExtra)
library(flexclust)
library(dbscan)
```

Use cluster analysis to explore and analyze the given dataset as follows:

#a. Use only the numerical variables (1 to 9) to cluster the 21 firms. Justify the various choices made in
conducting the cluster analysis, such as weights for different variables, the specific clustering algorithm(s)
used, the number of clusters formed, and so on

```{r}
Pharma <- read.csv("C:/Users/vaide/Downloads/Pharmaceuticals.csv")
head(Pharma)
```
#To remove any missing value that might be present in the data

```{r}
Pharma_data <- na.omit(Pharma)
```

#numerical variables from column 1 to 9 to cluster 21 firms

```{r}
row.names(Pharma_data)<- Pharma_data[,1]
Dataset<- Pharma_data[, 3:11]
head(Dataset)
```
#normalizing the data using Scale function

```{r}
Dataset2<- scale(Dataset)
head(Dataset2)
```

#Computing K-means clustering for different centers using multiple values of K and examine the differences in results

```{r}
k_m <- kmeans(Dataset2, centers = 2, nstart = 30)
k_m1<- kmeans(Dataset2, centers = 5, nstart = 30)
k_m2<- kmeans(Dataset2, centers = 6, nstart = 30)
P_1<-fviz_cluster(k_m, data = Dataset2)+ggtitle("k=2")
p_2<-fviz_cluster(k_m1, data = Dataset2)+ggtitle("k=5")
p_3<-fviz_cluster(k_m2, data = Dataset2)+ggtitle("k=6")
grid.arrange(P_1,p_2,p_3, nrow = 2)

```
#To get the best value of radius or eps.

#best value of radius at min points of 4.

```{r}
dbscan::kNNdistplot(Dataset, k=4)
```

# DBSCAN Algorithm at eps=30 and minpts =4

```{r}
D<- dbscan::dbscan(Dataset, eps = 30, minPts = 4)
print(D)
```

# DBSCAN Algorithm at eps=30 and minpts =4

```{r}
print(D$cluster)
```

# Visualization of clusters

```{r}
fviz_cluster(D, Dataset) + ggtitle("DBSCAN Plot")
```

#Choosing K-means over DBSCAN was motivated by the fact that K-means clustering can provide insights into the financial profiles of pharmaceutical businesses and that it is widely used in exploratory data analysis to identify patterns and groupings in the data. DBSCAN can be helpful in exposing groups of enterprises with similar financial features, which can aid in strategic decision-making and investment analysis for datasets with dense areas. It is also easy to understand.The K-means algorithm requires a fixed number of clusters, k. The user can decide how many clusters to construct, which may be useful in some situations. The number of clusters may not be easily determined using DBSCAN or hierarchical clustering.

```{r}

# euclidean distance
Eucl_d <- dist(Dataset2, method = "euclidean")

# Hierarchical Clustering
hclust <- hclust(Eucl_d, method = "complete")

# Visualize the output Dendrogram at height=5
plot(hclust, cex = 0.75, main = "Dendrogram of Hierarchical Clustering")
rect.hclust(hclust, h=5, border = 1:5)

```
```{r}
#Determining optimal clusters using Elbow method
dis <-  dist(Dataset2, method = "euclidean")

# Visualizing a distance matrix
fviz_dist(dis)
```
#Total within-cluster sum of squares (wss) for each k is tot.withinss, which is the total within-cluster sum of squares. 

```{r}
set.seed(123)
Total_wss<- function(k){
kmeans(Dataset2, k, nstart =10)$tot.withinss
}
k.values<- 1:10
Totalwss_cluster<- map_dbl(k.values, Total_wss)
plot(k.values, Totalwss_cluster,
     type="b", pch = 16, frame = TRUE,
     xlab="Number of clusters",
     ylab="Total within-clusters sum of squares")

```
#Though there is an elbow at position 2, it is still unclear because of the less precise graphic representation, as can be seen when looking at the above graph.

#Implication of Silhouette method

```{r}
fviz_nbclust(Dataset2,kmeans,method="silhouette")
```

#Because K=5 is clearly represented, the Silhouette approach was used for the final analysis. findings were extracted using five clusters, and the findings were visualized.

```{r}
set.seed(123)
fl<- kmeans(Dataset2, 5, nstart = 25)
print(fl)
fviz_cluster(fl, data = Dataset2)

```

#b. Interpret the clusters with respect to the numerical variables used in forming the clusters.

```{r}
Dataset%>%
  mutate(Cluster = fl$cluster) %>%
  group_by(Cluster)%>% summarise_all("mean")
  clusplot(Dataset2,fl$cluster, color = TRUE, labels = 2,lines = 0)
```

Cluster 1: AHM, SGP, WYE, BMY, AZN, ABT, NVS, LLY - This group has the lowest sales growth and the highest net profit margin. These businesses have slow revenue growth and comparatively little leverage. businesses have the best return on equity and net profit margin, indicating that businesses offer a wide range of profitable products. Thus, they don't have to use up all of their resources. These businesses have little leverage because they don't need to borrow money from the capital market.


Cluster 2: BAY, CHTT, IVX - This cluster has poor ROA, revenue growth, and net profit margin, but high leverage and beta. These businesses are examples of creative startups in the sector. Compared to those well-known brands, they are relatively minor in terms of market capitalization, and few people are familiar with their name. Their revenue growth is modest and their net profit margins are low due to their recent establishment, lack of experience, and lack of profitable products that can produce cash flow. They have a low ROA and a high degree of leverage since they heavily rely on R&D. Since they have a high beta and are investing in the future, their price will increase in a rising market.


Cluster 3: PHA and AGN There are just two businesses in this cluster: AGN and PHA. Its net profit margin, low ROA, lowest beta, and highest P/E ratio are all present.Consequently, many businesses had modest net profits in the past but great expectations for the future. The market values them highly since they may invest a significant amount of money in D&I in cutting-edge technologies. Nevertheless, investors bear greater risk due to its high price.


Cluster 4: The group with the largest market capitalization, high ROE and ROA, high net profit margin, high asset turnover, and low leverage. It includes JNJ, MRK, PFE, and GSK. These businesses represent the industry leaders with their highest market capitalizations and most noticeable positions. These businesses have the lowest leverage values, high ROE, ROA, asset turnover, and outstanding capital usage. With every dollar invested in these companies, they stand to benefit the most. In addition to mature products that require little capital or asset investment from the companies but generate substantial revenue and strong net profit margins, they must have a few best-selling and dominant items in the market. 

Cluster 5: AVE, MRX, WPI, and ELAN  This cluster features low market capitalization, low P/E, low turnover rate, high beta, and strong revenue growth. These conventional small-sized enterprises appear to lack significant capital utilization abilities based on their poor ROE, ROA, and turnover rates. Nevertheless, we can presume that the organizations are being led in the right way by either internal reformation or external market developments given the robust rate of revenue growth. Furthermore, the lowest P/E suggests that their share price is still modest.


#c. Is there a pattern in the clusters with respect to the numerical variables (10 to 12)? (those not used informing the clusters)

```{r}
pattern <- Pharma[12:14] %>% mutate(Clusters=fl$cluster)
ggplot(pattern, mapping = aes(factor(Clusters), fill =Median_Recommendation))+geom_bar(position='dodge')+labs(x ='Clusters',y ='Frequence')
ggplot(pattern, mapping = aes(factor(Clusters),fill = Location))+
  geom_bar(position = 'dodge')+labs(x ='Clusters',y = 'Frequence')
ggplot(pattern, mapping = aes(factor(Clusters),fill = Exchange))+geom_bar(position = 'dodge')+
  labs(x ='Clusters',y = 'Frequence')

```


#Group 1:

The median recommendation is that Cluster 1 has a solid hold.

Cluster 1 has three places, with the United States having a greater number than the other two. 
Switzerland and the Kingdom.

Exchange: The NYSE, the sole exchange in Cluster 1, is home to a vast array of 
attendees.


#Group 2:
Cluster 2 has a poor buy rating and a strong hold rating, according to the median recommendation.

Location: The US is ranked higher than Germany in two places within Cluster 2.

Exchange: AMEX, NASDAQ, and NYSE are the three exchanges located in Cluster 2. 
dispersed uniformly.


#Group 3:
The median recommendation is that Cluster 3 has a low purchase and a low hold.

Location: The United States and Canada are the only two places in Cluster 3 that are equally
dispersed.

Exchange: The NYSE is the only exchange in Cluster 3.

#Group 4:
The median recommendation is that Cluster 4 be bought and held at a high price. 

Location: There are two locations for Cluster 4, with the US having a significant advantage over the UK.

Exchange: The NYSE is the only exchange in Cluster 4.

#Group 5:

Cluster 5 includes a moderate buy and a moderate sell, per the median recommendation advice.

Location: There are three locations for Cluster 5, with the United States being the most notable.

Exchange: The NYSE is the sole exchange that is a part of Cluster 5.

 
#d.Provide an appropriate name for each cluster using any or all of the variables in the dataset


Cluster 1=Strong Hold
Cluster 2=poor purchase rate 
Cluster 3=low hold
Cluster 4=high price
Cluster 5=moderate sell






